A Road Segment Prioritization Approach for Cycling Infrastructure

### 2023-09-23 paper_drafts/paper_round_2.Rmd
edits for submission 3

geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach. The approach, which can be applied to other cities, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
output:
   # bookdown::github_document2:
#    bookdown::pdf_document2:
  bookdown::word_document2:
      highlight: null   # to remove code highlighting in r chunks
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
      # https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
      # pandoc_args:
      # - '--lua-filter=extras/scholarly-metadata.lua'
      # - '--lua-filter=extras/author-info-blocks.lua'
      keep_tex:  true
editor_options:
  markdown:
    wrap: sentence
# elsevier style for referencing (from https://github.com/citation-style-language/styles)
csl: extras/elsevier-harvard.csl
always_allow_html: yes
---

@@

```{r, include=FALSE}
library(tidyverse)
library(tmap)
knitr::opts_chunk$set(warning = FALSE)
```

@@

Funding alone is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
In this paper we propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
We used the city of Manchester as a case study.
The city is situated in the region of Greater Manchester, which has ambitions to become the first UK region with a carbon neutral transport network.
These ambitions are supported by plans for a 1800 mile walking and cycling network (as part of the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020].
While schemes have already been put in place for around a quarter of this network, the majority is yet to be built.

@@

Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion *within communities*.
We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
We propose three sub-methods that address some of the limitations of previous studies.
These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
The current work addresses these limitations through the following organisation.
Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'.
An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).

@@

## Case study area

Manchester is a city in the North West of England, and is the most populous of the 10 metropolitan boroughs that make up Greater Manchester (\@ref(fig:manchestermap)).
"Manchester" can be used to refer to multiple things:

-   Manchester City Council, which is a local government district that includes the city centre and surrounding areas.
-   The Greater Manchester region, which is a metropolitan county that includes the city of Manchester and 9 other local government districts.
-   The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas. This definition is formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics [@ons2015cities].

We use the TCITY definition of Manchester, which was "developed specifically for the production and analysis of statistics" and which is shown in Figure \@ref(fig:manchestermap).
Manchester is a cosmopolitan city with a large university population.
As shown in Figure \@ref(fig:manchestermap) (right), the high density of the central area allows relatively high levels of cycling compared to the rest of the city.

```{r manchestermap, echo=FALSE, fig.cap="Maps of Greater Manchester (left) and the case study area - City of Manchester (right)", fig.show='hold', fig.align="center"}
greater_manchester = pct::get_pct_zones(region = "greater-manchester", geography = "msoa")
greater_manchester = greater_manchester |>
  mutate(pcycle = bicycle / all * 100)
# names(greater_manchester)
msoas = sf::read_sf("./data/Manchester/msoas_geometry.shp")
msoas_union = sf::st_union(msoas)
greater_manchester_lads = greater_manchester |>
  group_by(lad_name) |>
  summarise()
?tm_scale_bar
m1 = tm_shape(greater_manchester_lads) +
  tm_polygons(col = "lad_name", palette = "Set1", alpha = 0.2, title = "") +
  tm_layout(legend.position = c("left", "bottom")) +
  tm_shape(msoas_union) +
  tm_borders(col = "black", lwd = 2) +
  tm_scale_bar(breaks = (0:3)*5)
msoas_centroids = sf::st_centroid(msoas)
pct_zones_in_manchester = greater_manchester[msoas_centroids, ] 
m2 = tm_shape(greater_manchester, bb = msoas) +
  tm_fill("pcycle", palette = "Blues", title = "% cycling to work\n(2011 Census)") +
  tm_layout(legend.position = c("left", "bottom"), legend.bg.color = "white", legend.bg.alpha = 0.5) +
  tm_shape(msoas_union) +
  tm_borders(col = "black", lwd = 2) +
  tm_scale_bar(breaks = 0:5)
tmap_arrange(m1, m2, nrow = 1)
```

In the context of other cities defined in the TCITY dataset, Manchester is a large city, with a mid-level of cycling (Figure \@ref(fig:manchestercyclinglevels)).
If the ambitions of the Bee Network are realized in the coming years, this mode share could go up significantly.

```{r manchestercyclinglevels, echo=FALSE, fig.cap="Cycling levels in cities across the country", out.width="85%", fig.show='hold', fig.align="center"}

knitr::include_graphics("./data/uk_cities_mode_share_2021.png")
```

>>>>

geometry: margin=2cm
bibliography:
  - mybibfile.bib
  - rlbibfile.bib
#linenumbers: true
numbersections: true
abstract: "Understanding the motivators and deterrents to cycling is essential for creating infrastructure that gets more people to adopt cycling as a mode of transport. This paper demonstrates a new approach to support the prioritization of cycling infrastructure and cycling network design, accounting for cyclist preferences and the growing emphasis on 'filtered permeability' and 'Low Traffic Neighborhood' interventions internationally. The approach combines distance decay, route calculation, and network analysis methods to examine where future cycling demand is most likely to arise, how such demand could be accommodated within existing street networks, and how to ensure a fair distribution of investment. Although each of these methods has been applied to cycling infrastructure prioritization in previous research, this is the first time that they have been combined, creating an integrated road segment prioritization approach. The approach, which can be applied to other cities, is demonstrated in a case study of Manchester, resulting in cycling networks that balance directness against the need for safe and stress-free routes under different investment scenarios. A key benefit of the approach from a policy perspective is its ability to support egalitarian and cost-effective strategic cycle network planning. \\par\\textbf{Keywords:} cycling networks, low-traffic neighborhoods, routing, transport equity"
# keywords: "cycling networks, low-traffic neighborhoods, potential demand, routing"
output:
 # bookdown::github_document2:
   bookdown::pdf_document2:
 # bookdown::word_document2:
      highlight: null   # to remove code highlighting in r chunks
      toc: false
      extra_dependencies: ["subcaption", "longtable", "float", "multirow", "placeins", "multicol"]
      # https://stackoverflow.com/questions/52918716/authors-and-affiliations-in-the-yaml-of-rmarkdown
      # pandoc_args:
      # - '--lua-filter=extras/scholarly-metadata.lua'
      # - '--lua-filter=extras/author-info-blocks.lua'
      keep_tex:  true
editor_options:
  markdown:
    wrap: sentence
# elsevier style for referencing (from https://github.com/citation-style-language/styles)
csl: extras/elsevier-harvard.csl
always_allow_html: yes
---

@@

@@

Funding alone is therefore no guarantee of a change in commuting across the country; it must be used to design adequate cycling infrastructure that is based on motivators and deterrents to cycling.
In this paper we propose an approach for planning connected cycling networks that can be generalised to any city with appropriate input data (as shown in the Appendix).
We used the city of Manchester as a case study due to its ambitions to become the first UK region with a carbon neutral transport network, supported by ambitious plans for a 1800 mile walking and cycling network (the 'Bee Network') that will be delivered over 10 years [@manchestercyclingbee2020].
While schemes have already been put in place for around a quarter of this network, the majority is yet to be built.

@@

Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion *within communities*.
We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
We propose three sub-methods that address some of the limitations of previous studies.
These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
The current work addresses these limitations through the following organisation.
Section \@ref(calculating-potential-cycling-demand) focuses on calculating potential cycling demand.
Section \@ref(routing) focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
Section \@ref(road-segment-prioritization) outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
It then introduces the network expansion algorithms, and compares an approach grounded in 'egalitarianism' to one grounded in 'utilitarianism'. 
As mentioned above, Manchester is used as a case study. 
An overview of the steps involved is shown in Figure \@ref(fig:methodflowchart).

@@

### 2023-09-23 paper_round_2.tex
edits for submission 3

Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion \emph{within communities}.
We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
We propose three sub-methods that address some of the limitations of previous studies.
These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
The current work addresses these limitations through the following organisation.
Section \ref{calculating-potential-cycling-demand} focuses on calculating potential cycling demand.
Section \ref{routing} focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
Section \ref{road-segment-prioritization} outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'.

As mentioned above, Manchester is used as a case study.
An overview of the steps involved is shown in Figure \ref{fig:methodflowchart}.

@@

\hypertarget{case-study-area}{%
\subsection{Case study area}\label{case-study-area}}

Manchester is a city in the North West of England.
``Manchester'' can be used to refer to multiple things:

\begin{itemize}
\tightlist
\item
  Manchester City Council, which is a local government district that includes the city centre and surrounding areas.
\item
  The Greater Manchester region, which is a metropolitan county that includes the city of Manchester and 9 other local government districts.
\item
  The city of Manchester, which is used by visitors to refer to the city centre and surrounding areas. This definition if formalised in the Major Towns and Cities (TCITY) dataset provided by the Office for National Statistics.
\end{itemize}

We use the TCITY definition of Manchester, which was ``developed specifically for the production and analysis of statistics'' and which is shown in Figure \ref{fig:manchestermap}.

\begin{verbatim}
##   [1] "geo_code"                "geo_name"               
##   [3] "lad11cd"                 "lad_name"               
##   [5] "all"                     "bicycle"                
##   [7] "foot"                    "car_driver"             
##   [9] "car_passenger"           "motorbike"              
##  [11] "train_tube"              "bus"                    
##  [13] "taxi_other"              "govtarget_slc"          
##  [15] "govtarget_sic"           "govtarget_slw"          
##  [17] "govtarget_siw"           "govtarget_sld"          
##  [19] "govtarget_sid"           "govtarget_slp"          
##  [21] "govtarget_sip"           "govtarget_slm"          
##  [23] "govtarget_sim"           "govtarget_slpt"         
##  [25] "govtarget_sipt"          "govnearmkt_slc"         
##  [27] "govnearmkt_sic"          "govnearmkt_slw"         
##  [29] "govnearmkt_siw"          "govnearmkt_sld"         
##  [31] "govnearmkt_sid"          "govnearmkt_slp"         
##  [33] "govnearmkt_sip"          "govnearmkt_slm"         
##  [35] "govnearmkt_sim"          "govnearmkt_slpt"        
##  [37] "govnearmkt_sipt"         "gendereq_slc"           
##  [39] "gendereq_sic"            "gendereq_slw"           
##  [41] "gendereq_siw"            "gendereq_sld"           
##  [43] "gendereq_sid"            "gendereq_slp"           
##  [45] "gendereq_sip"            "gendereq_slm"           
##  [47] "gendereq_sim"            "gendereq_slpt"          
##  [49] "gendereq_sipt"           "dutch_slc"              
##  [51] "dutch_sic"               "dutch_slw"              
##  [53] "dutch_siw"               "dutch_sld"              
##  [55] "dutch_sid"               "dutch_slp"              
##  [57] "dutch_sip"               "dutch_slm"              
##  [59] "dutch_sim"               "dutch_slpt"             
##  [61] "dutch_sipt"              "ebike_slc"              
##  [63] "ebike_sic"               "ebike_slw"              
##  [65] "ebike_siw"               "ebike_sld"              
##  [67] "ebike_sid"               "ebike_slp"              
##  [69] "ebike_sip"               "ebike_slm"              
##  [71] "ebike_sim"               "ebike_slpt"             
##  [73] "ebike_sipt"              "base_slcyclehours"      
##  [75] "govtarget_sicyclehours"  "govnearmkt_sicyclehours"
##  [77] "gendereq_sicyclehours"   "dutch_sicyclehours"     
##  [79] "ebike_sicyclehours"      "base_sldeath"           
##  [81] "base_slyll"              "base_slvalueyll"        
##  [83] "base_slsickdays"         "base_slvaluesick"       
##  [85] "base_slvaluecomb"        "govtarget_sideath"      
##  [87] "govtarget_siyll"         "govtarget_sivalueyll"   
##  [89] "govtarget_sisickdays"    "govtarget_sivaluesick"  
##  [91] "govtarget_sivaluecomb"   "govnearmkt_sideath"     
##  [93] "govnearmkt_siyll"        "govnearmkt_sivalueyll"  
##  [95] "govnearmkt_sisickdays"   "govnearmkt_sivaluesick" 
##  [97] "govnearmkt_sivaluecomb"  "gendereq_sideath"       
##  [99] "gendereq_siyll"          "gendereq_sivalueyll"    
## [101] "gendereq_sisickdays"     "gendereq_sivaluesick"   
## [103] "gendereq_sivaluecomb"    "dutch_sideath"          
## [105] "dutch_siyll"             "dutch_sivalueyll"       
## [107] "dutch_sisickdays"        "dutch_sivaluesick"      
## [109] "dutch_sivaluecomb"       "ebike_sideath"          
## [111] "ebike_siyll"             "ebike_sivalueyll"       
## [113] "ebike_sisickdays"        "ebike_sivaluesick"      
## [115] "ebike_sivaluecomb"       "base_slcarkm"           
## [117] "base_slco2"              "govtarget_sicarkm"      
## [119] "govtarget_sico2"         "govnearmkt_sicarkm"     
## [121] "govnearmkt_sico2"        "gendereq_sicarkm"       
## [123] "gendereq_sico2"          "dutch_sicarkm"          
## [125] "dutch_sico2"             "ebike_sicarkm"          
## [127] "ebike_sico2"             "perc_rf_dist_u10km"     
## [129] "avslope_perc_u10km"      "geometry"               
## [131] "pcycle"
\end{verbatim}

\begin{verbatim}
## Warning: st_centroid assumes attributes are constant over geometries
\end{verbatim}

\includegraphics{README_files/figure-latex/unnamed-chunk-4-1.pdf}

>>>>

Our work builds on these community finding approaches by proposing a similar greedy network expansion algorithm for cycle network expansion \emph{within communities}.
We incorporate community finding methods for study area partitioning with weighted routing to avoid links that are stressful to cycle on.
In doing so, we propose an approach that ensures a better distribution investments across the study area while also accounting for motivators and deterrents to cycling.
We propose three sub-methods that address some of the limitations of previous studies.
These limitations include (a) an inherent bias when basing network design solely on existing cycling demand, (b) the proposal of routes that may not correspond to studies on cyclist preference and government policies, and (c) an insufficient consideration of the ethical principles underlying the analysis.
The current work addresses these limitations through the following organisation.
Section \ref{calculating-potential-cycling-demand} focuses on calculating potential cycling demand.
Section \ref{routing} focuses on routing the demand onto the road network while accounting for cyclist preferences and government priorities.
Section \ref{road-segment-prioritization} outlines a method for partitioning the study area based on a community finding algorithm and routed cycling demand.
It then introduces the network expansion algorithms, and compares an approach grounded in `egalitarianism' to one grounded in `utilitarianism'.
As mentioned above, Manchester is used as a case study.
An overview of the steps involved is shown in Figure \ref{fig:methodflowchart}.

@@
